---
title: "GAMs Landela block"
author: Michael Wellington, Petra Kuhnert, Roger Lawes, Luigi Renzullo, Jamie Pittock,
  Peter Ramshaw, Martin Moyo, Emmanuel Kimaro, Miguel Tafula, Andre van Rooyen
date: '2022-08-22'
output:
  html_document:
    df_print: paged
vignette: |
  %\VignetteIndexEntry{GAMs Landela} 
  %\VignetteEncoding{UTF-8} 
  %\VignetteEngine{knitr::rmarkdown}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(mgcv)
library(tidyverse)
library(lubridate)
library(mgcViz)
library(viridis)
library(gridExtra)
library(ggpubr)
library(cowplot)
library(gratia)
library(gganimate)
library(gifski)
library(zoo)
library(ggsci)
library(ggbreak)
library(rgdal)
library(scales)
library(sf)
library(itsadug)
library(TISAschemeWUE)
```

## Bring in and clean data

We are using pre-prepared data from Landsat-7 and Landsat-8. We will select observations > 0 for GPP, E, T, and ET, and also filter to years greater than 2012. This is because Landsat-8 is required to give sufficient temporal resolution for trend analysis, and was launched in 2013.

```{r data, echo=FALSE}
data("Landela_df")

Landela_df <- Landela_df %>% 
  mutate_at(vars(date), dplyr::funs(year = lubridate::year, month = lubridate::month, day =lubridate::day))


# Clean data
Landela_df_nona <- Landela_df[!is.na(Landela_df$GPP) & (Landela_df$GPP > 0) & 
                                      !is.na(Landela_df$ETa) & (Landela_df$ETa > 0) &
                                      !is.na(Landela_df$Ea) & (Landela_df$Ea > 0) & 
                                      !is.na(Landela_df$Ta) & (Landela_df$Ta > 0) &
                                      (Landela_df$year > 2012),]
```

## Define GAM structure

Next, we define GAM formulae with different lags for rainfall. We will loop through these GAMs to select the best fit. We ended up choosing a 90 day fit as it delivered the lowest residual deviance across many, but not all, schemes. We kept this lag constant 

```{r GAMs}
# List of formulae with different time lags for rainfall (30, 60, 90, 120 day)
GPP_f_30 <- log(GPP) ~ 
  # smooth term for year
  s(year, bs="cr", k=9) + 
  # cyclic term for season
  s(month, bs="cc", k=12) +  
  # smooth term for spatial interaction
  s(x,y, k=50, bs='gp') +                  
  # seasonal within year
  ti(month, year, bs = c("cc", "cr"), k = c(12,9)) +  
  # space x time
  ti(x, y, year, d = c(2, 1), bs = c("gp", "cr"), 
     k = c(50,9)) +  
  # random effect for ls
  #s(sensor, bs = 're') + 
  # rainfall lag
  s(chirps_30, bs="cr", k=30) 

GPP_f_60 <- log(GPP) ~ 
  # smooth term for year
  s(year, bs="cr", k=9) + 
  # cyclic term for season
  s(month, bs="cc", k=12) +  
  # smooth term for spatial interaction
  s(x,y, k=50, bs='gp') +                  
  # seasonal within year
  ti(month, year, bs = c("cc", "cr"), k = c(12,9)) +  
  # space x time
  ti(x, y, year, d = c(2, 1), bs = c("gp", "cr"), 
     k = c(50,9)) +  
  # random effect for ls
  #s(sensor, bs = 're') + 
  # rainfall lag
  s(chirps_60, bs="cr", k=30) 

GPP_f_90 <- log(GPP) ~ 
  # smooth term for year
  s(year, bs="cr", k=9) + 
  # cyclic term for season
  s(month, bs="cc", k=12) +  
  # smooth term for spatial interaction
  s(x,y, k=50, bs='gp') +                  
  # seasonal within year
  ti(month, year, bs = c("cc", "cr"), k = c(12,9)) +  
  # space x time
  ti(x, y, year, d = c(2, 1), bs = c("gp", "cr"), 
     k = c(50,9)) +  
  # random effect for ls
  #s(sensor, bs = 're') + 
  # rainfall lag
  s(chirps_90, bs="cr", k=30) 

GPP_f_120 <- log(GPP) ~ 
  # smooth term for year
  s(year, bs="cr", k=9) + 
  # cyclic term for season
  s(month, bs="cc", k=12) +  
  # smooth term for spatial interaction
  s(x,y, k=50, bs='gp') +                  
  # seasonal within year
  ti(month, year, bs = c("cc", "cr"), k = c(12,9)) +  
  # space x time
  ti(x, y, year, d = c(2, 1), bs = c("gp", "cr"), 
     k = c(50,9)) +  
  # random effect for ls
  #s(sensor, bs = 're') + 
  # rainfall lag
  s(chirps_120, bs="cr", k=30)

GPP_fs <- list(GPP_f_30, GPP_f_60, GPP_f_90, GPP_f_120)

#Loop through different lags and test with anova
GPP_Landela_GAM1 <- bam(GPP_f_30, discrete=TRUE, nthreads=8, data=Landela_df_nona)
GPP_Landela_GAM2 <- bam(GPP_f_60, discrete=TRUE, nthreads=8, data=Landela_df_nona)
GPP_Landela_GAM3 <- bam(GPP_f_90, discrete=TRUE, nthreads=8, data=Landela_df_nona)
GPP_Landela_GAM4 <- bam(GPP_f_120, discrete=TRUE, nthreads=8, data=Landela_df_nona)

anova(GPP_Landela_GAM1, GPP_Landela_GAM2, GPP_Landela_GAM3, GPP_Landela_GAM4, test='F')
```

Now we will run GAMs on GPP, E,T, and ET. This code chunk is quite long so is hidden in this vignette, but can be found on the GitHub repository.

``` {r runGAMS, include=FALSE, echo=FALSE}

# Run GAMs for biomass (GPP)
GPP_Landela_GAM <- GPP_Landela_GAM3

resids_Landela_GPP <- residuals.gam(GPP_Landela_GAM)
valRho_Landela_GPP <- acf(resids_Landela_GPP, plot=FALSE)$acf[2]
GPP_Landela_GAM <- bam(GPP_f_90, discrete=TRUE, nthreads=8, data=Landela_df_nona, rho = valRho_Landela_GPP)
GPP_Landela_GAM1_vis <- getViz(GPP_Landela_GAM)

check(GPP_Landela_GAM1_vis)

GPP_Landela_year_plot <- plot_smooth(GPP_Landela_GAM, view = 'year', transform = exp)

##
GPP_Landela_year_plot_gg <- ggplot(data=GPP_Landela_year_plot$fv, aes(x=year, y=fit))+
  geom_line()+geom_ribbon(aes(ymin=ll, ymax=ul), alpha=0.2)+scale_x_continuous(breaks=seq(2013,2021,1))+
  labs(x="Year", y=expression(paste("GPP gC/m"^2,paste("/day"))))+theme_minimal()

GPP_Landela_space_plot <- plot(sm(GPP_Landela_GAM1_vis, 3), trans = function(x){ 
  GPP_Landela_GAM$family$linkinv(coef(GPP_Landela_GAM)[1] + x) 
}) 
GPP_Landela_space_plot_gg <- ggplot(data=(GPP_Landela_space_plot$data$fit%>% na.omit()), aes(x=x, y=y, z=tz))+
  geom_raster(aes(fill=tz)) + geom_contour(colour="black") + scale_fill_viridis_c(name=expression(paste("GPP gC/m"^2)))+
  theme(element_blank(), axis.text = element_blank()) +coord_equal() 

##

GPP_Landela_yearxspace <- plotSlice(sm(GPP_Landela_GAM1_vis,5), fix=list("year"=seq(2013,2021)), trans = exp)+labs(title = "GPP") +
  guides(fill=guide_colourbar("GPP"))

GPP_Landela_yearxspace_plot <- ggplot()+geom_raster(data=GPP_Landela_yearxspace$data$fit, aes(x=x, y=y, fill=tz), alpha=0.8)+
  scale_fill_viridis_c(labels=label_wrap(5))+theme_void()+
  facet_wrap(~.fx.year)+geom_contour(data=GPP_Landela_yearxspace$data$fit, aes(x=x, y=y, z=tz, col='red'))+
  labs(fill="GPP (gC/m2)")+coord_equal()+guides(alpha="none", colour="none")

GPP_Landela_yearxmonth <- plot(sm(GPP_Landela_GAM1_vis, 4), trans = function(x){ 
  GPP_Landela_GAM$family$linkinv(coef(GPP_Landela_GAM)[1] + x) 
}) 

GPP_Landela_yearxmonth_plot <- ggplot(data=GPP_Landela_yearxmonth$data$fit, aes(x=y, y=x, fill=tz)) +
  geom_raster()+
  geom_contour(aes(x=y, y=x, z=tz, col="red"), show.legend = FALSE)+
  scale_fill_viridis_c(labels=label_wrap(5))+
  scale_y_continuous(breaks=seq(2013, 2021, by=1))+
  scale_x_continuous(breaks=seq(1, 12, by=1))+
  labs(y="Year", x="Month", fill="GPP (gC/m2)") + theme_bw()

# Run GAMs for WUE, evapotranspiration (ETa), evaporation (Ea), transpiration (Ta) 
# Stick with 90 day lag

Landela_df_nona$WUE <- Landela_df_nona$GPP / Landela_df_nona$ETa

WUE_f_90 <- log(WUE) ~ 
  # smooth term for year
  s(year, bs="cr", k=9) + 
  # cyclic term for season
  s(month, bs="cc", k=12) +  
  # smooth term for spatial interaction
  s(x,y) +                  
  # seasonal within year
  ti(month, year, bs = c("cc", "cr"), k = c(12,9)) +  
  # space x time
  ti(x, y, year, d = c(2, 1), bs = c("tp", "cr"), 
     k = c(50,9)) +  
  # random effect for ls
  #s(sensor, bs = 're') + 
  # rainfall lag
  s(chirps_90, bs="cr", k=30)

ETa_f_90 <- log(ETa) ~ 
  # smooth term for year
  s(year, bs="cr", k=9) + 
  # cyclic term for season
  s(month, bs="cc", k=12) +  
  # smooth term for spatial interaction
  s(x,y) +                  
  # seasonal within year
  ti(month, year, bs = c("cc", "cr"), k = c(12,9)) +  
  # space x time
  ti(x, y, year, d = c(2, 1), bs = c("tp", "cr"), 
     k = c(50,9)) +  
  # random effect for ls
  #s(sensor, bs = 're') + 
  # rainfall lag
  s(chirps_90, bs="cr", k=30)

Ta_f_90 <- log(Ta) ~ 
  # smooth term for year
  s(year, bs="cr", k=9) + 
  # cyclic term for season
  s(month, bs="cc", k=12) +  
  # smooth term for spatial interaction
  s(x,y) +                  
  # seasonal within year
  ti(month, year, bs = c("cc", "cr"), k = c(12,9)) +  
  # space x time
  ti(x, y, year, d = c(2, 1), bs = c("tp", "cr"), 
     k = c(50,9)) +  
  # random effect for ls
  #s(sensor, bs = 're') + 
  # rainfall lag
  s(chirps_90, bs="cr", k=30)

Ea_f_90 <- log(Ea) ~ 
  # smooth term for year
  s(year, bs="cr", k=9) + 
  # cyclic term for season
  s(month, bs="cc", k=12) +  
  # smooth term for spatial interaction
  s(x,y) +                  
  # seasonal within year
  ti(month, year, bs = c("cc", "cr"), k = c(12,9)) +  
  # space x time
  ti(x, y, year, d = c(2, 1), bs = c("tp", "cr"), 
     k = c(50,9)) +  
  # random effect for ls
  #s(sensor, bs = 're') + 
  # rainfall lag
  s(chirps_90, bs="cr", k=30)

WUE_Landela_GAM <- bam(WUE_f_90, discrete=TRUE, nthreads=8, data = Landela_df_nona)
resids_Landela_WUE <- residuals.gam(WUE_Landela_GAM)
valRho_Landela_WUE <- acf(resids_Landela_WUE , plot=FALSE)$acf[2]
WUE_Landela_GAM <- bam(WUE_f_90, discrete=TRUE, nthreads=8, data = Landela_df_nona, rho = valRho_Landela_WUE)

WUE_Landela_GAM1_vis <- getViz(WUE_Landela_GAM)

ETa_Landela_GAM <- bam(ETa_f_90, discrete=TRUE, nthreads=8, data = Landela_df_nona)
resids_Landela_ETa <- residuals.gam(ETa_Landela_GAM)
valRho_Landela_ETa <- acf(resids_Landela_ETa , plot=FALSE)$acf[2]
ETa_Landela_GAM <- bam(ETa_f_90, discrete=TRUE, nthreads=8, data = Landela_df_nona, rho = valRho_Landela_ETa)

ETa_Landela_GAM1_vis <- getViz(ETa_Landela_GAM)

Ta_Landela_GAM <- bam(Ta_f_90, discrete=TRUE, nthreads=8, data = Landela_df_nona)
resids_Landela_Ta <- residuals.gam(Ta_Landela_GAM)
valRho_Landela_Ta <- acf(resids_Landela_Ta , plot=FALSE)$acf[2]
Ta_Landela_GAM <- bam(Ta_f_90, discrete=TRUE, nthreads=8, data = Landela_df_nona, rho = valRho_Landela_Ta)

Ta_Landela_GAM1_vis <- getViz(Ta_Landela_GAM)

Ea_Landela_GAM <- bam(Ea_f_90, discrete=TRUE, nthreads=8, data = Landela_df_nona)
resids_Landela_Ea <- residuals.gam(Ea_Landela_GAM)
valRho_Landela_Ea <- acf(resids_Landela_Ea , plot=FALSE)$acf[2]
Ea_Landela_GAM <- bam(Ea_f_90, discrete=TRUE, nthreads=8, data = Landela_df_nona, rho = valRho_Landela_Ea)

Ea_Landela_GAM1_vis <- getViz(Ea_Landela_GAM)

WUE_Landela_year_plot <- plot_smooth(WUE_Landela_GAM, view = 'year', transform = exp)

WUE_Landela_year_plot_gg <- ggplot(data=WUE_Landela_year_plot$fv, aes(x=year, y=fit))+
  geom_line()+geom_ribbon(aes(ymin=ll, ymax=ul), alpha=0.2)+scale_x_continuous(breaks=seq(2013,2021,1))+
  labs(x="Year", y="WUE gC/mm")+theme_minimal()

##
WUE_Landela_space_plot <- plot(sm(WUE_Landela_GAM1_vis, 3), trans = function(x){ 
  WUE_Landela_GAM$family$linkinv(coef(WUE_Landela_GAM)[1] + x) 
})
WUE_Landela_space_plot_gg <- ggplot(data=(WUE_Landela_space_plot$data$fit %>% na.omit()), aes(x=x, y=y, z=tz))+
  geom_raster(aes(fill=tz)) + geom_contour(colour="black") + scale_fill_viridis_c(name="WUE gC/mm")+
  theme(element_blank(), axis.text = element_blank()) +coord_equal()
##

ETa_Landela_year_plot <- plot_smooth(ETa_Landela_GAM, view = 'year', transform = exp)

##
ETa_Landela_space_plot <- plot(sm(ETa_Landela_GAM1_vis, 3), trans = function(x){ 
  ETa_Landela_GAM$family$linkinv(coef(ETa_Landela_GAM)[1] + x) 
})
ETa_Landela_space_plot_gg <- ggplot((data=ETa_Landela_space_plot$data$fit %>% na.omit()),
                                    aes(x=x, y=y, z=z))+
  geom_raster(aes(fill=z)) + geom_contour(colour="black") + 
  scale_fill_viridis_c(name="ETa (mm)", limits=c(-0.4,0.4), oob = scales::squish)+
  theme(element_blank(), axis.text = element_blank()) + coord_equal() 

ETa_Landela_yearxspace <- plotSlice(sm(ETa_Landela_GAM1_vis,5), fix=list("year"=seq(2013,2021)), trans = exp)+labs(title = "GPP") +
  guides(fill=guide_colourbar("GPP"))

ETa_Landela_yearxspace_plot <- ggplot()+geom_raster(data=ETa_Landela_yearxspace$data$fit, aes(x=x, y=y, fill=tz), alpha=0.8)+
  scale_fill_viridis_c(labels=label_wrap(5))+theme_void()+
  facet_wrap(~.fx.year)+geom_contour(data=ETa_Landela_yearxspace$data$fit, aes(x=x, y=y, z=tz, col='red'))+
  labs(fill="ETa mm/day")+coord_equal()+guides(alpha="none", colour="none") 
  

##

Ea_Landela_year_plot <- plot_smooth(Ea_Landela_GAM, view = 'year', transform = exp)

##
Ea_Landela_space_plot <- plot(sm(Ea_Landela_GAM1_vis, 3), trans = function(x){ 
  Ea_Landela_GAM$family$linkinv(coef(Ea_Landela_GAM)[1] + x) 
})
Ea_Landela_space_plot_gg <- ggplot(data=(Ea_Landela_space_plot$data$fit %>%  na.omit()), aes(x=x, y=y, z=z))+
  geom_raster(aes(fill=z)) + geom_contour(colour="black") + 
  scale_fill_viridis_c(name="Ea (mm)", limits=c(-0.4,0.4), oob = scales::squish)+
  theme(element_blank(), axis.text = element_blank())  + coord_equal()
##


Ta_Landela_year_plot <- plot_smooth(Ta_Landela_GAM, view = 'year', transform = exp)


##
Ta_Landela_space_plot <- plot(sm(Ta_Landela_GAM1_vis, 3), trans = function(x){ 
  Ea_Landela_GAM$family$linkinv(coef(Ea_Landela_GAM)[1] + x) 
})
Ta_Landela_space_plot_gg <- ggplot(data=(Ta_Landela_space_plot$data$fit %>% na.omit()), aes(x=x, y=y, z=z))+
  geom_raster(aes(fill=z)) + geom_contour(colour="black") + 
    scale_fill_viridis_c(name="Ta (mm)", limits=c(-0.4,0.4), oob = scales::squish)+
  theme(element_blank(), axis.text = element_blank()) + coord_equal()
##

GPP_Landela_plot_dat <- data.frame(var="GPP", value=GPP_Landela_year_plot$fv$fit,
                                      ll=GPP_Landela_year_plot$fv$ll,
                                      ul=GPP_Landela_year_plot$fv$ul,
                                      year=GPP_Landela_year_plot$fv$year)

WUE_Landela_plot_dat <- data.frame(var="WUE", value=WUE_Landela_year_plot$fv$fit,
                                      ll=WUE_Landela_year_plot$fv$ll,
                                      ul=WUE_Landela_year_plot$fv$ul,
                                      year=WUE_Landela_year_plot$fv$year)

E_Landela_plot_dat <- data.frame(var="E", value=Ea_Landela_year_plot$fv$fit,
                                    ll=Ea_Landela_year_plot$fv$ll,
                                    ul=Ea_Landela_year_plot$fv$ul,
                                    year=Ea_Landela_year_plot$fv$year)
T_Landela_plot_dat <- data.frame(var="T", value=Ta_Landela_year_plot$fv$fit,
                                    ll=Ta_Landela_year_plot$fv$ll,
                                    ul=Ta_Landela_year_plot$fv$ul,
                                    year=Ta_Landela_year_plot$fv$year)
ET_Landela_plot_dat <- data.frame(var="ET", value=ETa_Landela_year_plot$fv$fit,
                                     ll=ETa_Landela_year_plot$fv$ll,
                                     ul=ETa_Landela_year_plot$fv$ul,
                                     year=ETa_Landela_year_plot$fv$year)
All_Landela_plot_dat <- rbind(GPP_Landela_plot_dat, WUE_Landela_plot_dat,
                                 E_Landela_plot_dat, T_Landela_plot_dat, ET_Landela_plot_dat)

All_Landela_plot_dat$var <- as.factor(All_Landela_plot_dat$var)
All_Landela_plot_dat$var <- factor(All_Landela_plot_dat$var, levels=(c("WUE", "GPP", "ET", "E", "T")))
All_Landela_plot_dat$scheme <- "Landela"

ET_Landela_plot_dat <- rbind(E_Landela_plot_dat, T_Landela_plot_dat, ET_Landela_plot_dat)

ET_Landela_plot_dat$var <- as.factor(ET_Landela_plot_dat$var)
ET_Landela_plot_dat$var <- factor(ET_Landela_plot_dat$var, levels=(c("ET", "E", "T")))

ET_Landela_year_plot <- ggplot(data= ET_Landela_plot_dat, aes(x=year)) + 
  geom_line(aes(y=value, col=var))+
  geom_ribbon(aes(ymin=ll, ymax=ul, fill=var), alpha=0.2)+
  scale_x_continuous(breaks=seq(2013, 2021, by=1))+
  scale_color_manual(values=c("blue", "red", "darkgreen"))+
  scale_fill_manual(values=c("blue", "red", "darkgreen"))+
  theme(legend.title = element_blank())+
  labs(x="Year", y="mm/day")+
  theme_minimal() +theme(legend.title=element_blank())
ET_Landela_year_plot

##
laya <- rbind(c(1,2,3))
gs <- list(ETa_Landela_space_plot_gg, 
           Ta_Landela_space_plot_gg, Ea_Landela_space_plot_gg)
grid.arrange(grobs=gs, layout=laya, ncol=3)

ETa_Landela_oneplot_dat <- data.frame(x=ETa_Landela_space_plot$data$fit$x,
                                         y=ETa_Landela_space_plot$data$fit$y,
                                         z=ETa_Landela_space_plot$data$fit$z,
                                         var=c("Evapotranspiration")) %>% na.omit()

Ta_Landela_oneplot_dat <- data.frame(x=Ta_Landela_space_plot$data$fit$x,
                                        y=Ta_Landela_space_plot$data$fit$y,
                                        z=Ta_Landela_space_plot$data$fit$z,
                                        var=c("Transpiration")) %>% na.omit()

Ea_Landela_oneplot_dat <- data.frame(x=Ea_Landela_space_plot$data$fit$x,
                                        y=Ea_Landela_space_plot$data$fit$y,
                                        z=Ea_Landela_space_plot$data$fit$z,
                                        var=c("Evaporation")) %>% na.omit()

ET_Landela_oneplot_dat <- rbind(ETa_Landela_oneplot_dat, Ta_Landela_oneplot_dat,
                                   Ea_Landela_oneplot_dat)

ET_Landela_oneplot_dat$var <- factor(ET_Landela_oneplot_dat$var,
                                        levels=c("Evapotranspiration","Evaporation","Transpiration"))
ET_Landela_oneplot <- ggplot(data=ET_Landela_oneplot_dat, aes(x=x, y=y))+geom_raster(aes(fill=z))+
  geom_contour(aes(z=z), colour="black") + facet_grid(rows=vars(var)) + coord_equal() +
  scale_fill_viridis_c(name="mm", limits=c(-0.4,0.4), oob = scales::squish)+theme(element_blank(), axis.text = element_blank())
ET_Landela_oneplot

GPP_Landela_space_plot_dat <- as.data.frame(GPP_Landela_space_plot$data$fit) %>% na.omit()
GPP_Landela_space_plot_dat$var <- c("Gross Primary Productivity")

GPP_Landela_space_plot_gg <- ggplot(data=GPP_Landela_space_plot_dat, aes(x=x, y=y, z=tz))+
  geom_raster(aes(fill=tz)) + geom_contour(colour="black") + scale_fill_viridis_c(name=expression(paste("GPP gC/m"^2)))+
  facet_grid(rows=vars(var)) +
  theme(element_blank(), axis.text = element_blank()) +coord_equal() 

GPP_ET_Landela_space <- ggdraw() +
  draw_plot(GPP_Landela_space_plot_gg, x=0.5, y=0.3,width = 0.5, height=0.35)+
  draw_plot(ET_Landela_oneplot, x=0, y=0,width = 0.5, height=1)#+




##

p <- ggdraw() +
  draw_plot(WUE_Landela_year_plot_gg, x=0, y=0.8, width=0.9, height=0.2) +
  draw_plot(GPP_Landela_year_plot_gg, x=0, y=0.6, width=0.9, height=0.2) +
  #draw_plot(GPP_Landela_space_plot_gg, x=0.6, y=0.6,width = 0.4, height=0.2)+
  #draw_plot(WUE_Landela_space_plot_gg, x=0.6, y=0.8,width = 0.4, height=0.2)+
  draw_plot(ET_Landela_year_plot, x=0, y=0,width = 1, height=0.6)#+
#draw_plot(ET_Landela_oneplot, x=0.6, y=0,width = 0.4, height=0.6) 
```

## Plots

We can plot the annual and spatial trend results.
```{r plots}
Landela_GAMplot <- plot_grid(p, ncol=1, rel_heights=c(0.1, 1))
Landela_GAMplot
GPP_ET_Landela_space

```

## Linear GAMs

We need to change the GAM definition to extract linear trends for year, as below.

``` {r linearGAMs}
WUE_f_90_lin <- log(WUE) ~ 
  # smooth term for year
  year + 
  # cyclic term for season
  s(month, bs="cc", k=12) +  
  # smooth term for spatial interaction
  s(x,y) +                  
  # seasonal within year
  ti(month, year, bs = c("cc", "cr"), k = c(12,9)) +  
  # space x time
  ti(x, y, year, d = c(2, 1), bs = c("tp", "cr"), 
     k = c(50,9)) +  
  # rainfall lag
  s(chirps_90, bs="cr", k=30)

GPP_f_90_lin <- log(GPP) ~ 
  # smooth term for year
  year + 
  # cyclic term for season
  s(month, bs="cc", k=12) +  
  # smooth term for spatial interaction
  s(x,y, k=50, bs='gp') +                  
  # seasonal within year
  ti(month, year, bs = c("cc", "cr"), k = c(12,9)) +  
  # space x time
  ti(x, y, year, d = c(2, 1), bs = c("gp", "cr"), 
     k = c(50,9)) +  
  # rainfall lag
  s(chirps_90, bs="cr", k=30) 

ETa_f_90_lin <- log(ETa) ~ 
  # smooth term for year
  year + 
  # cyclic term for season
  s(month, bs="cc", k=12) +  
  # smooth term for spatial interaction
  s(x,y) +                  
  # seasonal within year
  ti(month, year, bs = c("cc", "cr"), k = c(12,9)) +  
  # space x time
  ti(x, y, year, d = c(2, 1), bs = c("tp", "cr"), 
     k = c(50,9)) +  
  # rainfall lag
  s(chirps_90, bs="cr", k=30)

Ta_f_90_lin <- log(Ta) ~ 
  # smooth term for year
  year + 
  # cyclic term for season
  s(month, bs="cc", k=12) +  
  # smooth term for spatial interaction
  s(x,y) +                  
  # seasonal within year
  ti(month, year, bs = c("cc", "cr"), k = c(12,9)) +  
  # space x time
  ti(x, y, year, d = c(2, 1), bs = c("tp", "cr"), 
     k = c(50,9)) +  
  # rainfall lag
  s(chirps_90, bs="cr", k=30)

Ea_f_90_lin <- log(Ea) ~ 
  # smooth term for year
  year + 
  # cyclic term for season
  s(month, bs="cc", k=12) +  
  # smooth term for spatial interaction
  s(x,y) +                  
  # seasonal within year
  ti(month, year, bs = c("cc", "cr"), k = c(12,9)) +  
  # space x time
  ti(x, y, year, d = c(2, 1), bs = c("tp", "cr"), 
     k = c(50,9)) +  
  # rainfall lag
  s(chirps_90, bs="cr", k=30)
```

Finally, we extract the % annual change, as below. We can see there is a 4.8% trend in GPP.

``` {r GPP_lin}

GPP_Landela_GAM1_lin <- bam(GPP_f_90_lin, discrete=TRUE, nthreads=8, data=Landela_df_nona)
(exp(GPP_Landela_GAM1_lin$coefficients["year"]) - 1)*100

```

Similarly, there is a 4.6% trend in T.

``` {r T_lin}

Ta_Landela_GAM1_lin <- bam(Ta_f_90_lin, discrete=TRUE, nthreads=8, data=Landela_df_nona)
(exp(Ta_Landela_GAM1_lin$coefficients["year"]) - 1)*100

```

Inversely, there is a -13.5% trend in E.

``` {r E_lin}

Ea_Landela_GAM1_lin <- bam(Ea_f_90_lin, discrete=TRUE, nthreads=8, data=Landela_df_nona)
(exp(Ea_Landela_GAM1_lin$coefficients["year"]) - 1)*100

```

Finally, there is a -3.6% trend in ET, indicating that GPP was decoupled from ET over this period at Landela.

``` {r ET_lin}
ETa_Landela_GAM1_lin <- bam(ETa_f_90_lin, discrete=TRUE, nthreads=8, data=Landela_df_nona)
(exp(ETa_Landela_GAM1_lin$coefficients["year"]) - 1)*100
```