---
title: "GAMs 25 de Setembro"
author: Michael Wellington, Petra Kuhnert, Roger Lawes, Luigi Renzullo, Jamie Pittock,
  Peter Ramshaw, Martin Moyo, Emmanuel Kimaro, Miguel Tafula, Andre van Rooyen
date: '2022-08-22'
output:
  html_document:
    df_print: paged
vignette: |
  %\VignetteIndexEntry{GAMs 25 de Setembro} 
  %\VignetteEncoding{UTF-8} 
  %\VignetteEngine{knitr::rmarkdown}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(mgcv)
library(tidyverse)
library(lubridate)
library(mgcViz)
library(viridis)
library(gridExtra)
library(ggpubr)
library(cowplot)
library(gratia)
library(gganimate)
library(gifski)
library(zoo)
library(ggsci)
library(ggbreak)
library(rgdal)
library(scales)
library(sf)
library(itsadug)
library(TISAschemeWUE)


```

## Bring in and clean data

We are using pre-prepared data from Landsat-7 and Landsat-8. We will select observations > 0 for GPP, E, T, and ET, and also filter to years greater than 2012. This is because Landsat-8 is required to give sufficient temporal resolution for trend analysis, and was launched in 2013.

```{r data, include=FALSE}
library(TISAschemeWUE)
data("deSetembro_df")

deSetembro_df <- deSetembro_df %>% 
  mutate_at(vars(date), dplyr::funs(year = lubridate::year, month = lubridate::month, day =lubridate::day))


# Clean data
deSetembro_df_nona <- deSetembro_df[!is.na(deSetembro_df$GPP) & (deSetembro_df$GPP > 0) & 
                                      !is.na(deSetembro_df$ETa) & (deSetembro_df$ETa > 0) &
                                      !is.na(deSetembro_df$Ea) & (deSetembro_df$Ea > 0) & 
                                      !is.na(deSetembro_df$Ta) & (deSetembro_df$Ta > 0) &
                                      (deSetembro_df$year > 2012),]
```

## Define GAM structure

Next, we define GAM formulae with different lags for rainfall. We will loop through these GAMs to select the best fit. We ended up choosing a 90 day fit as it delivered the lowest residual deviance across many, but not all, schemes. We kept this lag constant 

```{r GAMs}
# List of formulae with different time lags for rainfall (30, 60, 90, 120 day)
GPP_f_30 <- log(GPP) ~ 
  # smooth term for year
  s(year, bs="cr", k=9) + 
  # cyclic term for season
  s(month, bs="cc", k=12) +  
  # smooth term for spatial interaction
  s(x,y, k=50, bs='gp') +                  
  # seasonal within year
  ti(month, year, bs = c("cc", "cr"), k = c(12,9)) +  
  # space x time
  ti(x, y, year, d = c(2, 1), bs = c("gp", "cr"), 
     k = c(50,9)) +  
  # rainfall lag
  s(chirps_30, bs="cr", k=30) 

GPP_f_60 <- log(GPP) ~ 
  # smooth term for year
  s(year, bs="cr", k=9) + 
  # cyclic term for season
  s(month, bs="cc", k=12) +  
  # smooth term for spatial interaction
  s(x,y, k=50, bs='gp') +                  
  # seasonal within year
  ti(month, year, bs = c("cc", "cr"), k = c(12,9)) +  
  # space x time
  ti(x, y, year, d = c(2, 1), bs = c("gp", "cr"), 
     k = c(50,9)) +  
  # rainfall lag
  s(chirps_60, bs="cr", k=30) 

GPP_f_90 <- log(GPP) ~ 
  # smooth term for year
  s(year, bs="cr", k=9) + 
  # cyclic term for season
  s(month, bs="cc", k=12) +  
  # smooth term for spatial interaction
  s(x,y, k=50, bs='gp') +                  
  # seasonal within year
  ti(month, year, bs = c("cc", "cr"), k = c(12,9)) +  
  # space x time
  ti(x, y, year, d = c(2, 1), bs = c("gp", "cr"), 
     k = c(50,9)) +  
  # rainfall lag
  s(chirps_90, bs="cr", k=30) 

GPP_f_120 <- log(GPP) ~ 
  # smooth term for year
  s(year, bs="cr", k=9) + 
  # cyclic term for season
  s(month, bs="cc", k=12) +  
  # smooth term for spatial interaction
  s(x,y, k=50, bs='gp') +                  
  # seasonal within year
  ti(month, year, bs = c("cc", "cr"), k = c(12,9)) +  
  # space x time
  ti(x, y, year, d = c(2, 1), bs = c("gp", "cr"), 
     k = c(50,9)) +  
  # rainfall lag
  s(chirps_120, bs="cr", k=30)

GPP_fs <- list(GPP_f_30, GPP_f_60, GPP_f_90, GPP_f_120)

#Loop through different lags and test with anova
GPP_deSetembro_GAM1 <- bam(GPP_f_30, discrete=TRUE, nthreads=8, data=deSetembro_df_nona)
GPP_deSetembro_GAM2 <- bam(GPP_f_60, discrete=TRUE, nthreads=8, data=deSetembro_df_nona)
GPP_deSetembro_GAM3 <- bam(GPP_f_90, discrete=TRUE, nthreads=8, data=deSetembro_df_nona)
GPP_deSetembro_GAM4 <- bam(GPP_f_120, discrete=TRUE, nthreads=8, data=deSetembro_df_nona)

anova(GPP_deSetembro_GAM1, GPP_deSetembro_GAM2, GPP_deSetembro_GAM3, GPP_deSetembro_GAM4, test='F')
```

## Run GAMs

Now we will run GAMs on GPP, E,T, and ET. This code chunk is quite long so is hidden in this vignette, but can be found on the GitHub repository.

``` {r runGAMSGPP, include=FALSE, echo=FALSE}

# Run GAMs for biomass (GPP)
GPP_deSetembro_GAM <- GPP_deSetembro_GAM3

resids_deSetembro_GPP <- residuals.gam(GPP_deSetembro_GAM)
valRho_deSetembro_GPP <- acf(resids_deSetembro_GPP, plot=FALSE)$acf[2]
GPP_deSetembro_GAM <- bam(GPP_f_90, discrete=TRUE, nthreads=8, data=deSetembro_df_nona, rho = valRho_deSetembro_GPP)
GPP_deSetembro_GAM1_vis <- getViz(GPP_deSetembro_GAM)

check(GPP_deSetembro_GAM1_vis)

GPP_deSetembro_year_plot <- plot_smooth(GPP_deSetembro_GAM, view = 'year', transform = exp)

##
GPP_deSetembro_year_plot_gg <- ggplot(data=GPP_deSetembro_year_plot$fv, aes(x=year, y=fit))+
  geom_line()+geom_ribbon(aes(ymin=ll, ymax=ul), alpha=0.2)+scale_x_continuous(breaks=seq(2013,2021,1))+
  labs(x="Year", y=expression(paste("GPP gC/m"^2,paste("/day"))))+theme_minimal()

GPP_deSetembro_space_plot <- plot(sm(GPP_deSetembro_GAM1_vis, 3), trans = function(x){ 
  GPP_deSetembro_GAM$family$linkinv(coef(GPP_deSetembro_GAM)[1] + x) 
}) 
GPP_deSetembro_space_plot_gg <- ggplot(data=(GPP_deSetembro_space_plot$data$fit%>% na.omit()), aes(x=x, y=y, z=tz))+
  geom_raster(aes(fill=tz)) + geom_contour(colour="black") + scale_fill_viridis_c(name=expression(paste("GPP gC/m"^2)))+
  theme(element_blank(), axis.text = element_blank()) +coord_equal() 

##

GPP_deSetembro_yearxspace <- plotSlice(sm(GPP_deSetembro_GAM1_vis,5), fix=list("year"=seq(2013,2021)), trans = exp)+labs(title = "GPP") +
  guides(fill=guide_colourbar("GPP"))

GPP_deSetembro_yearxspace_plot <- ggplot()+geom_raster(data=GPP_deSetembro_yearxspace$data$fit, aes(x=x, y=y, fill=tz), alpha=0.8)+
  scale_fill_viridis_c(labels=label_wrap(5))+theme_void()+
  facet_wrap(~.fx.year)+geom_contour(data=GPP_deSetembro_yearxspace$data$fit, aes(x=x, y=y, z=tz, col='red'))+
  labs(fill="GPP (gC/m2)")+coord_equal()+guides(alpha="none", colour="none")

GPP_deSetembro_yearxmonth <- plot(sm(GPP_deSetembro_GAM1_vis, 4), trans = function(x){ 
  GPP_deSetembro_GAM$family$linkinv(coef(GPP_deSetembro_GAM)[1] + x) 
}) 

GPP_deSetembro_yearxmonth_plot <- ggplot(data=GPP_deSetembro_yearxmonth$data$fit, aes(x=y, y=x, fill=tz)) +
  geom_raster()+
  geom_contour(aes(x=y, y=x, z=tz, col="red"), show.legend = FALSE)+
  scale_fill_viridis_c(labels=label_wrap(5))+
  scale_y_continuous(breaks=seq(2013, 2021, by=1))+
  scale_x_continuous(breaks=seq(1, 12, by=1))+
  labs(y="Year", x="Month", fill="GPP (gC/m2)") + theme_bw()

# Run GAMs for WUE, evapotranspiration (ETa), evaporation (Ea), transpiration (Ta) 
# Stick with 90 day lag

deSetembro_df_nona$WUE <- deSetembro_df_nona$GPP / deSetembro_df_nona$ETa

WUE_f_90 <- log(WUE) ~ 
  # smooth term for year
  s(year, bs="cr", k=9) + 
  # cyclic term for season
  s(month, bs="cc", k=12) +  
  # smooth term for spatial interaction
  s(x,y) +                  
  # seasonal within year
  ti(month, year, bs = c("cc", "cr"), k = c(12,9)) +  
  # space x time
  ti(x, y, year, d = c(2, 1), bs = c("tp", "cr"), 
     k = c(50,9)) +  
  # random effect for ls
  #s(sensor, bs = 're') + 
  # rainfall lag
  s(chirps_90, bs="cr", k=30)

ETa_f_90 <- log(ETa) ~ 
  # smooth term for year
  s(year, bs="cr", k=9) + 
  # cyclic term for season
  s(month, bs="cc", k=12) +  
  # smooth term for spatial interaction
  s(x,y) +                  
  # seasonal within year
  ti(month, year, bs = c("cc", "cr"), k = c(12,9)) +  
  # space x time
  ti(x, y, year, d = c(2, 1), bs = c("tp", "cr"), 
     k = c(50,9)) +  
  # random effect for ls
  #s(sensor, bs = 're') + 
  # rainfall lag
  s(chirps_90, bs="cr", k=30)

Ta_f_90 <- log(Ta) ~ 
  # smooth term for year
  s(year, bs="cr", k=9) + 
  # cyclic term for season
  s(month, bs="cc", k=12) +  
  # smooth term for spatial interaction
  s(x,y) +                  
  # seasonal within year
  ti(month, year, bs = c("cc", "cr"), k = c(12,9)) +  
  # space x time
  ti(x, y, year, d = c(2, 1), bs = c("tp", "cr"), 
     k = c(50,9)) +  
  # random effect for ls
  #s(sensor, bs = 're') + 
  # rainfall lag
  s(chirps_90, bs="cr", k=30)

Ea_f_90 <- log(Ea) ~ 
  # smooth term for year
  s(year, bs="cr", k=9) + 
  # cyclic term for season
  s(month, bs="cc", k=12) +  
  # smooth term for spatial interaction
  s(x,y) +                  
  # seasonal within year
  ti(month, year, bs = c("cc", "cr"), k = c(12,9)) +  
  # space x time
  ti(x, y, year, d = c(2, 1), bs = c("tp", "cr"), 
     k = c(50,9)) +  
  # random effect for ls
  #s(sensor, bs = 're') + 
  # rainfall lag
  s(chirps_90, bs="cr", k=30)

WUE_deSetembro_GAM <- bam(WUE_f_90, discrete=TRUE, nthreads=8, data = deSetembro_df_nona)
resids_deSetembro_WUE <- residuals.gam(WUE_deSetembro_GAM)
valRho_deSetembro_WUE <- acf(resids_deSetembro_WUE , plot=FALSE)$acf[2]
WUE_deSetembro_GAM <- bam(WUE_f_90, discrete=TRUE, nthreads=8, data = deSetembro_df_nona, rho = valRho_deSetembro_WUE)

WUE_deSetembro_GAM1_vis <- getViz(WUE_deSetembro_GAM)

ETa_deSetembro_GAM <- bam(ETa_f_90, discrete=TRUE, nthreads=8, data = deSetembro_df_nona)
resids_deSetembro_ETa <- residuals.gam(ETa_deSetembro_GAM)
valRho_deSetembro_ETa <- acf(resids_deSetembro_ETa , plot=FALSE)$acf[2]
ETa_deSetembro_GAM <- bam(ETa_f_90, discrete=TRUE, nthreads=8, data = deSetembro_df_nona, rho = valRho_deSetembro_ETa)

ETa_deSetembro_GAM1_vis <- getViz(ETa_deSetembro_GAM)

Ta_deSetembro_GAM <- bam(Ta_f_90, discrete=TRUE, nthreads=8, data = deSetembro_df_nona)
resids_deSetembro_Ta <- residuals.gam(Ta_deSetembro_GAM)
valRho_deSetembro_Ta <- acf(resids_deSetembro_Ta , plot=FALSE)$acf[2]
Ta_deSetembro_GAM <- bam(Ta_f_90, discrete=TRUE, nthreads=8, data = deSetembro_df_nona, rho = valRho_deSetembro_Ta)

Ta_deSetembro_GAM1_vis <- getViz(Ta_deSetembro_GAM)

Ea_deSetembro_GAM <- bam(Ea_f_90, discrete=TRUE, nthreads=8, data = deSetembro_df_nona)
resids_deSetembro_Ea <- residuals.gam(Ea_deSetembro_GAM)
valRho_deSetembro_Ea <- acf(resids_deSetembro_Ea , plot=FALSE)$acf[2]
Ea_deSetembro_GAM <- bam(Ea_f_90, discrete=TRUE, nthreads=8, data = deSetembro_df_nona, rho = valRho_deSetembro_Ea)

Ea_deSetembro_GAM1_vis <- getViz(Ea_deSetembro_GAM)

WUE_deSetembro_year_plot <- plot_smooth(WUE_deSetembro_GAM, view = 'year', transform = exp)

WUE_deSetembro_year_plot_gg <- ggplot(data=WUE_deSetembro_year_plot$fv, aes(x=year, y=fit))+
  geom_line()+geom_ribbon(aes(ymin=ll, ymax=ul), alpha=0.2)+scale_x_continuous(breaks=seq(2013,2021,1))+
  labs(x="Year", y="WUE gC/mm")+theme_minimal()

##
WUE_deSetembro_space_plot <- plot(sm(WUE_deSetembro_GAM1_vis, 3), trans = function(x){ 
  WUE_deSetembro_GAM$family$linkinv(coef(WUE_deSetembro_GAM)[1] + x) 
})
WUE_deSetembro_space_plot_gg <- ggplot(data=(WUE_deSetembro_space_plot$data$fit %>% na.omit()), aes(x=x, y=y, z=tz))+
  geom_raster(aes(fill=tz)) + geom_contour(colour="black") + scale_fill_viridis_c(name="WUE gC/mm")+
  theme(element_blank(), axis.text = element_blank()) +coord_equal()
##

ETa_deSetembro_year_plot <- plot_smooth(ETa_deSetembro_GAM, view = 'year', transform = exp)

##
ETa_deSetembro_space_plot <- plot(sm(ETa_deSetembro_GAM1_vis, 3), trans = function(x){ 
  ETa_deSetembro_GAM$family$linkinv(coef(ETa_deSetembro_GAM)[1] + x) 
})
ETa_deSetembro_space_plot_gg <- ggplot((data=ETa_deSetembro_space_plot$data$fit %>% na.omit()),
                                    aes(x=x, y=y, z=z))+
  geom_raster(aes(fill=z)) + geom_contour(colour="black") + 
  scale_fill_viridis_c(name="ETa (mm)", limits=c(-0.4,0.4), oob = scales::squish)+
  theme(element_blank(), axis.text = element_blank()) + coord_equal() 

ETa_deSetembro_yearxspace <- plotSlice(sm(ETa_deSetembro_GAM1_vis,5), fix=list("year"=seq(2013,2021)), trans = exp)+labs(title = "GPP") +
  guides(fill=guide_colourbar("GPP"))

ETa_deSetembro_yearxspace_plot <- ggplot()+geom_raster(data=ETa_deSetembro_yearxspace$data$fit, aes(x=x, y=y, fill=tz), alpha=0.8)+
  scale_fill_viridis_c(labels=label_wrap(5))+theme_void()+
  facet_wrap(~.fx.year)+geom_contour(data=ETa_deSetembro_yearxspace$data$fit, aes(x=x, y=y, z=tz, col='red'))+
  labs(fill="ETa mm/day")+coord_equal()+guides(alpha="none", colour="none") 
  

##

Ea_deSetembro_year_plot <- plot_smooth(Ea_deSetembro_GAM, view = 'year', transform = exp)

##
Ea_deSetembro_space_plot <- plot(sm(Ea_deSetembro_GAM1_vis, 3), trans = function(x){ 
  Ea_deSetembro_GAM$family$linkinv(coef(Ea_deSetembro_GAM)[1] + x) 
})
Ea_deSetembro_space_plot_gg <- ggplot(data=(Ea_deSetembro_space_plot$data$fit %>%  na.omit()), aes(x=x, y=y, z=z))+
  geom_raster(aes(fill=z)) + geom_contour(colour="black") + 
  scale_fill_viridis_c(name="Ea (mm)", limits=c(-0.4,0.4), oob = scales::squish)+
  theme(element_blank(), axis.text = element_blank())  + coord_equal()
##


Ta_deSetembro_year_plot <- plot_smooth(Ta_deSetembro_GAM, view = 'year', transform = exp)


##
Ta_deSetembro_space_plot <- plot(sm(Ta_deSetembro_GAM1_vis, 3), trans = function(x){ 
  Ea_deSetembro_GAM$family$linkinv(coef(Ea_deSetembro_GAM)[1] + x) 
})
Ta_deSetembro_space_plot_gg <- ggplot(data=(Ta_deSetembro_space_plot$data$fit %>% na.omit()), aes(x=x, y=y, z=z))+
  geom_raster(aes(fill=z)) + geom_contour(colour="black") + 
    scale_fill_viridis_c(name="Ta (mm)", limits=c(-0.4,0.4), oob = scales::squish)+
  theme(element_blank(), axis.text = element_blank()) + coord_equal()
##

GPP_deSetembro_plot_dat <- data.frame(var="GPP", value=GPP_deSetembro_year_plot$fv$fit,
                                      ll=GPP_deSetembro_year_plot$fv$ll,
                                      ul=GPP_deSetembro_year_plot$fv$ul,
                                      year=GPP_deSetembro_year_plot$fv$year)

WUE_deSetembro_plot_dat <- data.frame(var="WUE", value=WUE_deSetembro_year_plot$fv$fit,
                                      ll=WUE_deSetembro_year_plot$fv$ll,
                                      ul=WUE_deSetembro_year_plot$fv$ul,
                                      year=WUE_deSetembro_year_plot$fv$year)

E_deSetembro_plot_dat <- data.frame(var="E", value=Ea_deSetembro_year_plot$fv$fit,
                                    ll=Ea_deSetembro_year_plot$fv$ll,
                                    ul=Ea_deSetembro_year_plot$fv$ul,
                                    year=Ea_deSetembro_year_plot$fv$year)
T_deSetembro_plot_dat <- data.frame(var="T", value=Ta_deSetembro_year_plot$fv$fit,
                                    ll=Ta_deSetembro_year_plot$fv$ll,
                                    ul=Ta_deSetembro_year_plot$fv$ul,
                                    year=Ta_deSetembro_year_plot$fv$year)
ET_deSetembro_plot_dat <- data.frame(var="ET", value=ETa_deSetembro_year_plot$fv$fit,
                                     ll=ETa_deSetembro_year_plot$fv$ll,
                                     ul=ETa_deSetembro_year_plot$fv$ul,
                                     year=ETa_deSetembro_year_plot$fv$year)
All_deSetembro_plot_dat <- rbind(GPP_deSetembro_plot_dat, WUE_deSetembro_plot_dat,
                                 E_deSetembro_plot_dat, T_deSetembro_plot_dat, ET_deSetembro_plot_dat)

All_deSetembro_plot_dat$var <- as.factor(All_deSetembro_plot_dat$var)
All_deSetembro_plot_dat$var <- factor(All_deSetembro_plot_dat$var, levels=(c("WUE", "GPP", "ET", "E", "T")))
All_deSetembro_plot_dat$scheme <- "deSetembro"

ET_deSetembro_plot_dat <- rbind(E_deSetembro_plot_dat, T_deSetembro_plot_dat, ET_deSetembro_plot_dat)

ET_deSetembro_plot_dat$var <- as.factor(ET_deSetembro_plot_dat$var)
ET_deSetembro_plot_dat$var <- factor(ET_deSetembro_plot_dat$var, levels=(c("ET", "E", "T")))

ET_deSetembro_year_plot <- ggplot(data= ET_deSetembro_plot_dat, aes(x=year)) + 
  geom_line(aes(y=value, col=var))+
  geom_ribbon(aes(ymin=ll, ymax=ul, fill=var), alpha=0.2)+
  scale_x_continuous(breaks=seq(2013, 2021, by=1))+
  scale_color_manual(values=c("blue", "red", "darkgreen"))+
  scale_fill_manual(values=c("blue", "red", "darkgreen"))+
  theme(legend.title = element_blank())+
  labs(x="Year", y="mm/day")+
  theme_minimal() +theme(legend.title=element_blank())
ET_deSetembro_year_plot

##
laya <- rbind(c(1,2,3))
gs <- list(ETa_deSetembro_space_plot_gg, 
           Ta_deSetembro_space_plot_gg, Ea_deSetembro_space_plot_gg)
grid.arrange(grobs=gs, layout=laya, ncol=3)

ETa_deSetembro_oneplot_dat <- data.frame(x=ETa_deSetembro_space_plot$data$fit$x,
                                         y=ETa_deSetembro_space_plot$data$fit$y,
                                         z=ETa_deSetembro_space_plot$data$fit$z,
                                         var=c("Evapotranspiration")) %>% na.omit()

Ta_deSetembro_oneplot_dat <- data.frame(x=Ta_deSetembro_space_plot$data$fit$x,
                                        y=Ta_deSetembro_space_plot$data$fit$y,
                                        z=Ta_deSetembro_space_plot$data$fit$z,
                                        var=c("Transpiration")) %>% na.omit()

Ea_deSetembro_oneplot_dat <- data.frame(x=Ea_deSetembro_space_plot$data$fit$x,
                                        y=Ea_deSetembro_space_plot$data$fit$y,
                                        z=Ea_deSetembro_space_plot$data$fit$z,
                                        var=c("Evaporation")) %>% na.omit()

ET_deSetembro_oneplot_dat <- rbind(ETa_deSetembro_oneplot_dat, Ta_deSetembro_oneplot_dat,
                                   Ea_deSetembro_oneplot_dat)

ET_deSetembro_oneplot_dat$var <- factor(ET_deSetembro_oneplot_dat$var,
                                        levels=c("Evapotranspiration","Evaporation","Transpiration"))
ET_deSetembro_oneplot <- ggplot(data=ET_deSetembro_oneplot_dat, aes(x=x, y=y))+geom_raster(aes(fill=z))+
  geom_contour(aes(z=z), colour="black") + facet_grid(rows=vars(var)) + coord_equal() +
  scale_fill_viridis_c(name="mm", limits=c(-0.4,0.4), oob = scales::squish)+theme(element_blank(), axis.text = element_blank())
ET_deSetembro_oneplot

GPP_deSetembro_space_plot_dat <- as.data.frame(GPP_deSetembro_space_plot$data$fit) %>% na.omit()
GPP_deSetembro_space_plot_dat$var <- c("Gross Primary Productivity")

GPP_deSetembro_space_plot_gg <- ggplot(data=GPP_deSetembro_space_plot_dat, aes(x=x, y=y, z=tz))+
  geom_raster(aes(fill=tz)) + geom_contour(colour="black") + scale_fill_viridis_c(name=expression(paste("GPP gC/m"^2)))+
  facet_grid(rows=vars(var)) +
  theme(element_blank(), axis.text = element_blank()) +coord_equal() 

GPP_ET_deSetembro_space <- ggdraw() +
  draw_plot(GPP_deSetembro_space_plot_gg, x=0.5, y=0.3,width = 0.5, height=0.35)+
  draw_plot(ET_deSetembro_oneplot, x=0, y=0,width = 0.5, height=1)#+




##

p <- ggdraw() +
  draw_plot(WUE_deSetembro_year_plot_gg, x=0, y=0.8, width=0.9, height=0.2) +
  draw_plot(GPP_deSetembro_year_plot_gg, x=0, y=0.6, width=0.9, height=0.2) +
  #draw_plot(GPP_deSetembro_space_plot_gg, x=0.6, y=0.6,width = 0.4, height=0.2)+
  #draw_plot(WUE_deSetembro_space_plot_gg, x=0.6, y=0.8,width = 0.4, height=0.2)+
  draw_plot(ET_deSetembro_year_plot, x=0, y=0,width = 1, height=0.6)#+
#draw_plot(ET_deSetembro_oneplot, x=0.6, y=0,width = 0.4, height=0.6) 
```

## Plots

Finally, we can plot the annual and spatial trend results.
```{r plots}
deSetembro_GAMplot <- plot_grid(p, ncol=1, rel_heights=c(0.1, 1))
deSetembro_GAMplot
GPP_ET_deSetembro_space

```